# Data sources

In this project, we have three data sources

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Vaccination data
We collected vaccination data from an organization called 'Our World in Data'. They gather vaccination information of United States from official and update the data in a daily basis. The daily cases and death data are gathered from Center for Systems Science and Engineering at John Hopkins University. It's updated every 24 hours as well. 

For the vaccination data, there are 14 features. The names of them are listed below:
```{r}
print(colnames(state_vaccination_raw))
```

The starting date of this dataset is 2021-01-12 and the last day is yesterday.
The first problem we noticed was that in location, there are many places that are not US states. Like "Bureau of Prisons","Dept of Defense"... Therefore, the first step that we did was to filter out these none-states. Then we found a common issue, missing data. We visualized it in missing values chapter and decided to do a backward filling. Because most of these missing values were not continuous, we could simply use the latest values to fill in these missing spots.

For the cases and death data, the raw data put every day as a column therefore we first converted these columns to rows.
```{r}
print(colnames(state_cases_raw)[c(1:15)])
```

The starting date of this dataset is 2020-01-22 and the last day is yesterday.
When we tried to calculate daily growing cases and death from cummulative data, we noticed that there were mistakes when collecting these data because some of the growing numbers were negative. We used the same filling method as in vaccination data.



## COVID-19-Reported-Patient-Impact-and-Hospital-Capa

```{r,message=FALSE,warning=FALSE,echo=FALSE}
source('data/hospital_preprocessing.R')
```

This data-set provides facility-level data for hospital utilization aggregated on a weekly basis (Friday to Thursday) in the time period from 2020-07-31 to 2021-11-26. It is mainly collected from two sources: (1) HHS TeleTracking, and (2) reporting provided directly to HHS Protect by state/territorial health departments on behalf of their healthcare facilities.

There are 106 features with 309719 records in the original data-set, but not all of those features are useful for our analysis, so we only select 6 specific features from 106 original features. These selected features including date, state, capacities of hospitals and number of hospitalized patients who are confirmed with COVID-19. 

We also noticed that there is value -999999 in data-set which is not a reasonable value. Replacing -999999 with 2 since it represents sums and averages less than four, so use average to replace that. Because we want to compare the information between states, so group the data-set by state is also needed. Moreover, as the summary of the data-set shows below, there are missing values in features. Therefore, we also need to handle the NA values in the later process.

```{r}
#summary of data-set

summary(hospital_capacity_selected)

```

## United-States-COVID-19-Cases-and-Deaths-by-State


This data-set is collected by CDC, which includes counts of COVID-19 cases and death numbers daily in the time period form 2020-01-22 to 2021-12-07. CDC collects and organizes these data based on these most recent numbers reported by states, territories, and other jurisdictions. There are 41160 records with 15 features in the original data-set, and each record represents a specific date for a specific state. We only keep 3 features which are state, total number of cases and total number of death for our further analysis.

Because we want to merge this data-set with 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa', so it is important to make sure they are in the same time period. Since the time range for 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' is smaller, so the subset of original data-set is generated with the time period from 2020-07-31 to 2021-11-26. The sub data-set has 29040 records with 3 features. Note that for the features 'tot_cases' and 'tot_death', they are cumulative features during the time period. Moreover, to compare between state, we add population column to calculate the percentage.

```{r}
#summary of data-set

summary(case_death_selected)

```














