# Data transformation

## Vaccination, case, and death data
The code is in data/data_preprocessing.R

For this part, there are 3 datasets: vaccination, cases, and death. Cases and death are from same source, vaccination is from another source.

Vaccination:
1. Filter out non-state, only keep 50 states. 
2. Fill NA with previous value.
3. Rename 'location' to 'state' for later join with other dataframe. Change 'New York State' to 'New York' to be consistent with other data source.
4. When calculating growth, fill first date with 0 otherwise it would be NA.

Dataframe after transformation:
```{r}
head(state_vaccination)
```

Case and death:
1. Filter out non-state, only keep 50 states.
2. Convert date columns to rows by using dplyr.
3. Change date from character type to date type.
4. Rename 'Province_state' to 'state' for later join with other dataframe.
5. When calculating growth, fill first date with 0 otherwise it would be NA. Change all negative values to NA, then fill with previous value. There shouldn't have negative values in growth because cumulative covid cases and deaths shouldn't decrease as time goes by.

Dataframes after transformation:
```{r}
head(state_cases)

head(state_death)
```
















