# Data transformation

## Vaccination, case, and death data
The code is in data/data_preprocessing.R

For this part, there are 3 datasets: vaccination, cases, and death. Cases and death are from same source, vaccination is from another source.

Vaccination:
1. Filter out non-state, only keep 50 states. 
2. Fill NA with previous value.
3. Rename 'location' to 'state' for later join with other dataframe. Change 'New York State' to 'New York' to be consistent with other data source.
4. When calculating growth, fill first date with 0 otherwise it would be NA.

Dataframe after transformation:
```{r}
head(state_vaccination)
```

Case and death:
1. Filter out non-state, only keep 50 states.
2. Convert date columns to rows by using dplyr.
3. Change date from character type to date type.
4. Rename 'Province_state' to 'state' for later join with other dataframe.
5. When calculating growth, fill first date with 0 otherwise it would be NA. Change all negative values to NA, then fill with previous value. There shouldn't have negative values in growth because cumulative covid cases and deaths shouldn't decrease as time goes by.

Dataframes after transformation:
```{r}
head(state_cases)

head(state_death)
```

## Relationship between morbidity and mortality and hospital capacity
In order to see the relationship, two data-set 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' and 'United-States-COVID-19-Cases-and-Deaths-by-State' are used.

### COVID-19_Hospital_Capacity

Hospital capacity:
1. replace -999999 with 2
2. select useful features
3. rename columns with shorter column names 
4. fill NA by using mean value for each feature in each state
5. group all records for each hospital to calculate each hospital's average number of beds and icu beds in the time period
6. group all hospital records for each state to calculate the total confirmed COVID-19 cases in that state during the time period

```{r}
#source('data/hospital_preprocessing.R')
```

Dataframes after transformation:
```{r}
# fill NA by using mean value for each feature in each state
impute_mean <- function(x) replace(x, is.na(x), round(mean(x, na.rm = TRUE)))
hospital_capacity_selected_fill <- ddply(hospital_capacity_selected, ~ state, transform, total_beds = impute_mean(total_beds),
     total_icu_beds = impute_mean(total_icu_beds),adu_conf = impute_mean(adu_conf),ped_conf = impute_mean(ped_conf))
```

```{r}
hospital_area=hospital_capacity_selected_fill[!duplicated(hospital_capacity_selected_fill$hospital_pk), ]
hospital_area=hospital_area[,c('hospital_pk','state')]
hospital_each <- hospital_capacity_selected_fill %>% group_by(hospital_pk) %>% summarize_if(is.numeric,mean,na.rm = TRUE)
hospital_each=hospital_each[,c('hospital_pk','total_beds','total_icu_beds')]
hospital_each<-merge(hospital_each,hospital_area,by="hospital_pk")
hospital_each['total_beds']=round(hospital_each['total_beds'])
hospital_each['total_icu_beds']=round(hospital_each['total_icu_beds'])
hospital_each <- hospital_each %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)


hospital_state <- hospital_capacity_selected_fill %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)
hospital_state['total_beds']=hospital_each['total_beds']
hospital_state['total_icu_beds']=hospital_each['total_icu_beds']
summary(hospital_state)
```

After the transformation, there are 56 observations with 5 features, each record represents one state in the United States.


### COVID-19_Case_Death

Death case:
1. convert type of submission_date to date 
2. select date which is in the same range as COVID-19 hospital capacity data-set
3. select useful features
4. add column 'population' by state
5. keep the record with latest date for each state as features 'tot_cases' and 'tot_death' are cumulative features

Dataframes after transformation:
```{r}
case_state <- case_death_selected [case_death_selected $submission_date==end_date_hos,]
summary(case_state )
```

After the selection, there are 52 records with 5 features. Each record represents one state in the United States.

### Merge

In order to see the relationship between hospital capacity and death rate of COVID-19, added new columns which would represents the death rate, total number of confirmed inpatients, hospitalization rate and percentage of total beds in population for each state. 

```{r}
hospital_case<-merge(hospital_state,case_state,by="state")

# remove date column
hospital_case<-hospital_case[,-grep("submission_date",colnames(hospital_case))]

# calculate death rate for each state
hospital_case['death_rate']<-hospital_case$tot_death/hospital_case$tot_cases

# calculate total number of confirmed inpatient for each state
hospital_case['tot_confi_inpatient']<-hospital_case$adu_conf+hospital_case$ped_conf

# calculate the hospitalization rate for each state
hospital_case['inhospi_rate']<-hospital_case$tot_confi_inpatient/hospital_case$tot_cases

# calculate the percentage of number of beds in total population for each state
hospital_case['beds_perc']<-(hospital_case$total_beds+hospital_case$total_icu_beds)/hospital_case$population

summary(hospital_case)
```

So the final data-set after transformation will have 52 observations with 12 features.















