# Data transformation

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Vaccination, case, and death data
The code is in data/data_preprocessing.R

For this part, there are 3 datasets: vaccination, cases, and death. Cases and death are from same source, vaccination is from another source.

Vaccination:
1. Filter out non-state, only keep 50 states. 
2. Fill NA with previous value.
3. Rename 'location' to 'state' for later join with other dataframe. Change 'New York State' to 'New York' to be consistent with other data source.
4. When calculating growth, fill first date with 0 otherwise it would be NA.

Dataframe after transformation:
```{r}
head(state_vaccination)
```

Case and death:
1. Filter out non-state, only keep 50 states.
2. Convert date columns to rows by using dplyr.
3. Change date from character type to date type.
4. Rename 'Province_state' to 'state' for later join with other dataframe.
5. When calculating growth, fill first date with 0 otherwise it would be NA. Change all negative values to NA, then fill with previous value. There shouldn't have negative values in growth because cumulative covid cases and deaths shouldn't decrease as time goes by.

Dataframes after transformation:
```{r}
head(state_cases)

head(state_death)
```

## Relationship between morbidity and mortality and hospital capacity
In order to do see the relationship, two data-set 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' and 'United-States-COVID-19-Cases-and-Deaths-by-State' are used.

### COVID-19_Hospital_Capacity
To see the difference in hospital capacity and number of inpatient who is confirmed with COVID-19 between state, group all hospital records for each state. 

```{r}
source('data/hospital_preprocessing.R')
```

```{r}
# fill NA by using mean value for each feature in each state
impute_mean <- function(x) replace(x, is.na(x), round(mean(x, na.rm = TRUE)))
hospital_capacity_selected_fill <- ddply(hospital_capacity_selected, ~ state, transform, total_beds = impute_mean(total_beds),
     total_icu_beds = impute_mean(total_icu_beds),adu_conf = impute_mean(adu_conf),ped_conf = impute_mean(ped_conf))

hospital_state <- hospital_capacity_selected_fill %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)

summary(hospital_state)
```

After the transformation, there are 56 observations with 7 features, each record represents one state in the United States.


### COVID-19_Case_Death
To merge it with the Hospital Capacity data-set, records also needed to be grouped by state. However, as mentioned before, as features 'tot_cases' and 'tot_death', they are cumulative features during the time period, so only keep the record with latest date for each state. 

```{r}
case_state <- case_death_selected [case_death_selected $submission_date==end_date_hos,]
summary(case_state )
```

After the selection, there is 60 records with 4 features.

### Merge
In order to see the relationship between hospital capacity and death rate of COVID-19, added new columns which would represents the death rate, total number of confirmed inpatients, hospitalization rate, percentage of total beds in population and percentage of total icu beds in population for each state. 

```{r}
hospital_case<-merge(hospital_state,case_state,by="state")

# remove date column
hospital_case<-hospital_case[,-grep("submission_date",colnames(hospital_case))]

# calculate death rate for each state
hospital_case['death_rate']<-hospital_case$tot_death/hospital_case$tot_cases

# calculate total number of confirmed inpatient for each state
hospital_case['tot_confi_inpatient']<-hospital_case$adu_conf+hospital_case$ped_conf

# calculate the hospitalization rate for each state
hospital_case['inhospi_rate']<-hospital_case$tot_confi_inpatient/hospital_case$tot_cases

# calculate the percentage of number of beds in total population for each state
hospital_case['beds_perc']<-hospital_case$total_beds/hospital_case$population

# calculate the percentage of number of icu beds in total population for each state
hospital_case['icu_beds_perc']<-hospital_case$total_icu_beds/hospital_case$population

summary(hospital_case)
```


















