# Missing values

We will use US vaccination data as an example to show how we are going to show missing data pattern.

```{r,message=FALSE,warning=FALSE,echo=FALSE}
source('data_preprocessing.R')
```

## Dataframe after data cleaning
```{r}
head(us_whole_vaccination)
```

The original dataframe countain vaccination information of all states and some organizations of United States. We extracted whole country data and we can see that each row represents a day from 2020-12-20 to now. The information is updated everyday.

## Missing value counts by column
```{r,echo=FALSE}
aa <- colSums(is.na(us_whole_vaccination)) %>% sort(decreasing = TRUE)
result <- data.frame(num_missing = aa)
result
```

## Missing value counts by row
```{r,echo=FALSE}
values <- rowSums(is.na(us_whole_vaccination))
result <- data.frame(date=as.Date(character()),
                     num_missing=integer())
na_rows_index <- c()
count <- 1
for(i in 1:length(values)){
  if(values[i] != 0){
    result[count,1] <- us_whole_vaccination[i,1]
    result[count,2] <- values[i]
    count <- count + 1
    na_rows_index <- c(na_rows_index,i)
  }
}
result1 <- result
result1$date <- format(result1$date,"%Y-%m")
result1 <- result1 %>% group_by(date) %>% summarise(num_missing = sum(num_missing))
result1
```

Each row in the original dataframe represents a specific day. However, there are too many rows that have missing values. Therefore, we count missing values by month.
December 2020 and Janurary 2021 have the greatest number of missing values. This is reasonable as vaccination just started and it's hard to get all data well-organized.

## Heatmap
```{r,echo=FALSE,fig.height=10,fig.width=10}
df <- us_whole_vaccination[na_rows_index,-2] %>% 
  gather(key,value,-date) %>%
  mutate(missing=ifelse(is.na(value),'yes','no'))

ggplot(df,aes(x=key,y=fct_rev(as.character(date)),fill=missing)) +
  geom_tile(color='white') +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,vjust=0.6)) +
  labs(y='date')
  
```

```{r}

```









