# Results

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Latest cummulative cases and deaths
```{r}
day <- min(max(state_cases$date),max(state_death$date),max(state_vaccination$date))

ggplot(state_cases[state_cases$date==day,],aes(reorder(state,cases),cases)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Cases',title=paste('Total Cases On ',day)) +
  theme_bw()

ggplot(state_death[state_death$date==day,],aes(reorder(state,death),death)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Death',title=paste('Total Death On ',day)) +
  theme_bw()
```

The graphs above shows current confirmed cases and death of 50 states. We can see that California, Texas, Florida, New York, and Illinois are top 5 states with most confirmed cases. California, Texas, Florida, New York, and Pennsylvania are top 5 states with most deaths. 


## Latest vaccination situation
```{r,fig.width=9,fig.height=9,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases[,-4],by=c('date','state')),state_death[,-4])
df <- df[,c(-1,-2)]
cormat <- round(cor(df),2)

melted_cormat <- melt(cormat)

ggplot(data=melted_cormat,aes(Var1,Var2,fill=value))+
  geom_tile() +
  geom_text(aes(label=value)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = 'Correlation Heat Map of Vaccination Features')
```

As there are 12 features in our vaccination data, we use a correlation heatmap to select the most relevant features to confirmed cases and deaths. From the graph above we can see the top 4 features are total vaccinations, total distributed, people vaccinated, and people fully vaccinated. We will use people vaccinated and people fully vaccinated as two features to represent vaccination.

```{r}
df <- state_vaccination[,c(1,2,5,6)]
p1 <- ggplot(df[df$date == day,],aes(reorder(state, people_vaccinated),people_vaccinated)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Vaccinated') +
  theme_bw()

p2 <- ggplot(df[df$date == day,],aes(reorder(state, people_fully_vaccinated_per_hundred),people_fully_vaccinated_per_hundred)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Fully Vaccinated Per Hundred') +
  theme_bw()

p1 + p2 + plot_annotation(title=paste('Vaccination Situation On ',day))
```

The 5 states with most deaths also have the largest number of people vaccinated. However, none of them are on top of the list of people fully vaccinated.

## Time series of serious states
```{r,warning=FALSE,message=FALSE}
states <- c('California', 'Texas', 'Florida', 'New York', 'Pennsylvania')

df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(state %in% states) 
  
p1 <- ggplot(df, aes(x=date, y= case_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Case Growth')

p2 <- ggplot(df, aes(x=date, y= death_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Death Growth')

p3 <- ggplot(df, aes(x=date, y= vacc_diff)) +
  geom_line(aes(color=state),size=1) +
  theme_minimal() +
  labs(y='Vacc Growth')

p4 <- ggplot(df, aes(x=date, y= fully_vacc_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Fully Vacc Growth')

p1 + p2 + p3 + p4 + plot_layout(ncol = 1) + plot_annotation(title='Time Series')
```

There is a significant outlier in people_vaccinated daily growth graph. Because from October 2nd to November 28th, the people_vaccinated information in Pennsylvania is missing. We are doing a filling by pervious data, therefore the huge spike is false.

## Distribution
```{r,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(date == day)

ggplot(df, aes(cases, people_fully_vaccinated, color=state)) +
  geom_point(show.legend = FALSE) +
  geom_text(aes(label=state),size=2.6,hjust=0,vjust=0) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x='Cases',y='People Fully Vaccinated',title=paste('Vaccination vs. Cases Scatterplot On ',day))
  
```



## percentage of number of beds in population and death rate
```{r}
ggplot(hospital_case, aes(beds_perc, death_rate)) + 
  geom_point(size=1) +
  geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
  theme_bw() +
  ggtitle("Scatter Plot for beds VS death rate")

```
No clear trend occurs. It may indicates that there is no strong relationship between percentage of number of beds in population and death rate.


## percentage of number of icu beds in population and death rate

```{r}
ggplot(hospital_case, aes(icu_beds_perc, death_rate)) + 
  geom_point(size=1) +
  geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
  ggtitle("Scatter Plot for icu beds VS death rate")

```
There is a more clear trend between total number of icu hospital beds and death rate. As number of beds increase, there is an increase trend in the value of death rate.

## inpatient rate and death rate

```{r}
ggplot(hospital_case, aes(inhospi_rate, death_rate)) + 
  geom_point(size=1) +
  geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
  ggtitle("Scatter Plot for tinpatient rate VS death rate")

```

Death rate increase as the inpatient rate. Two possible explanations. The first one is people would not go to hospital unless they are in very bad situation. Thus as the number of impatient increase, death rate increase. The second explanation is consider about the time range. There is no very useful way to cure COVID-19 patient. Thus as inpatient number increase, death rate increase. 








