--- 
title: "United States Covid Project"
author: "Ruize Yu, Jiazhen Li, Wenxin Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
always_allow_html: yes
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

Beginning in 2020, COVID-19 has begun to spread all over the world. People equated it with the flu from the beginning, and later became more and more aware of the harm caused by this virus. Over time, although people have begun to develop vaccines to minimize the spread of COVID-19, COVID-19 is constantly mutating. Every day there are new infections and deaths.
<<<<<<< HEAD

Therefore, in this project, we would like to study some issues about COVID-19. Specifically, this is mainly divided into three parts. The first part is the impact of the vaccination and the spread of the epidemic. The second part mainly studies the relationship between hospital capacity and mortality.The last part of the study is.....

In the first part, analysis are based on the data collected from the United States. With knowledge that after vaccination started in United States, cases and death dropped significantly at the beginning, but after a while, a turning point appeared and now the cases are increasing just like before vaccination. The analysis will be done for the impact of Coronavirus vaccination on the prevalence and mortality. We will start from whole country and then break down to serious state to try to figure out why cases are increasing with huge number of vaccinations.


The second part uses data distributed by states in the United States. It mainly includes the number of hospital beds in each state, the diagnosis rate in a specific time period, the rate of consultations and the mortality rate. So it can be analysed that whether there is a relationship between hospital capacity and mortality, and the relationship between hospitalization and mortality.

The last part....

<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

In this project, we have three data sources

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Vaccination data
We collected vaccination data from an organization called 'Our World in Data'. They gather vaccination information of United States from official and update the data in a daily basis. The daily cases and death data are gathered from Center for Systems Science and Engineering at John Hopkins University. It's updated every 24 hours as well. 

For the vaccination data, there are 14 features. The names of them are listed below:
```{r}
print(colnames(state_vaccination_raw))
```

The starting date of this dataset is 2021-01-12 and the last day is yesterday.
The first problem we noticed was that in location, there are many places that are not US states. Like "Bureau of Prisons","Dept of Defense"... Therefore, the first step that we did was to filter out these none-states. Then we found a common issue, missing data. We visualized it in missing values chapter and decided to do a backward filling. Because most of these missing values were not continuous, we could simply use the latest values to fill in these missing spots.

For the cases and death data, the raw data put every day as a column therefore we first converted these columns to rows.
```{r}
print(colnames(state_cases_raw)[c(1:15)])
```

The starting date of this dataset is 2020-01-22 and the last day is yesterday.
When we tried to calculate daily growing cases and death from cummulative data, we noticed that there were mistakes when collecting these data because some of the growing numbers were negative. We used the same filling method as in vaccination data.

## COVID-19-Reported-Patient-Impact-and-Hospital-Capa

```{r,message=FALSE,warning=FALSE,echo=FALSE}
source('data/hospital_preprocessing.R')
```

This data-set provides facility-level data for hospital utilization aggregated on a weekly basis (Friday to Thursday) in the time period from 2020-07-31 to 2021-11-26. It is mainly collected from two sources: (1) HHS TeleTracking, and (2) reporting provided directly to HHS Protect by state/territorial health departments on behalf of their healthcare facilities.

There are 106 features with 309719 records in the original data-set, but not all of those features are useful for our analysis, so we only select 6 specific features from 106 original features. These selected features including date, state, capacities of hospitals and number of hospitalized patients who are confirmed with COVID-19. 

We also noticed that there is value -999999 in data-set which is not a reasonable value. Replacing -999999 with 2 since it represents sums and averages less than four, so use average to replace that. Because we want to compare the information between states, so group the data-set by state is also needed. Moreover, as the summary of the data-set shows below, there are missing values in features. Therefore, we also need to handle the NA values in the later process.

```{r}
#summary of data-set

summary(hospital_capacity_selected)

```

## United-States-COVID-19-Cases-and-Deaths-by-State


This data-set is collected by CDC, which includes counts of COVID-19 cases and death numbers daily in the time period form 2020-01-22 to 2021-12-07. CDC collects and organizes these data based on these most recent numbers reported by states, territories, and other jurisdictions. There are 41160 records with 15 features in the original data-set, and each record represents a specific date for a specific state. We only keep 3 features which are state, total number of cases and total number of death for our further analysis.

Because we want to merge this data-set with 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa', so it is important to make sure they are in the same time period. Since the time range for 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' is smaller, so the subset of original data-set is generated with the time period from 2020-07-31 to 2021-11-26. The sub data-set has 29040 records with 3 features. Note that for the features 'tot_cases' and 'tot_death', they are cumulative features during the time period. Moreover, to compare between state, we add population column to calculate the percentage.

```{r}
#summary of data-set

summary(case_death_selected)

```















<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

## Vaccination, case, and death data
The code is in data/data_preprocessing.R

For this part, there are 3 datasets: vaccination, cases, and death. Cases and death are from same source, vaccination is from another source.

Vaccination:
1. Filter out non-state, only keep 50 states. 
2. Fill NA with previous value.
3. Rename 'location' to 'state' for later join with other dataframe. Change 'New York State' to 'New York' to be consistent with other data source.
4. When calculating growth, fill first date with 0 otherwise it would be NA.

Dataframe after transformation:
```{r}
head(state_vaccination)
```

Case and death:
1. Filter out non-state, only keep 50 states.
2. Convert date columns to rows by using dplyr.
3. Change date from character type to date type.
4. Rename 'Province_state' to 'state' for later join with other dataframe.
5. When calculating growth, fill first date with 0 otherwise it would be NA. Change all negative values to NA, then fill with previous value. There shouldn't have negative values in growth because cumulative covid cases and deaths shouldn't decrease as time goes by.

Dataframes after transformation:
```{r}
head(state_cases)

head(state_death)
```

## Relationship between morbidity and mortality and hospital capacity
In order to see the relationship, two data-set 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' and 'United-States-COVID-19-Cases-and-Deaths-by-State' are used.

### COVID-19_Hospital_Capacity

Hospital capacity:
1. replace -999999 with 2
2. select useful features
3. rename columns with shorter column names 
4. fill NA by using mean value for each feature in each state
5. group all records for each hospital to calculate each hospital's average number of beds and icu beds in the time period
6. group all hospital records for each state to calculate the total confirmed COVID-19 cases in that state during the time period

```{r}
#source('data/hospital_preprocessing.R')
```

Dataframes after transformation:
```{r}
# fill NA by using mean value for each feature in each state
impute_mean <- function(x) replace(x, is.na(x), round(mean(x, na.rm = TRUE)))
hospital_capacity_selected_fill <- ddply(hospital_capacity_selected, ~ state, transform, total_beds = impute_mean(total_beds),
     total_icu_beds = impute_mean(total_icu_beds),adu_conf = impute_mean(adu_conf),ped_conf = impute_mean(ped_conf))
```

```{r}
hospital_area=hospital_capacity_selected_fill[!duplicated(hospital_capacity_selected_fill$hospital_pk), ]
hospital_area=hospital_area[,c('hospital_pk','state')]
hospital_each <- hospital_capacity_selected_fill %>% group_by(hospital_pk) %>% summarize_if(is.numeric,mean,na.rm = TRUE)
hospital_each=hospital_each[,c('hospital_pk','total_beds','total_icu_beds')]
hospital_each<-merge(hospital_each,hospital_area,by="hospital_pk")
hospital_each['total_beds']=round(hospital_each['total_beds'])
hospital_each['total_icu_beds']=round(hospital_each['total_icu_beds'])
hospital_each <- hospital_each %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)


hospital_state <- hospital_capacity_selected_fill %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)
hospital_state['total_beds']=hospital_each['total_beds']
hospital_state['total_icu_beds']=hospital_each['total_icu_beds']
summary(hospital_state)
```

After the transformation, there are 56 observations with 5 features, each record represents one state in the United States.


### COVID-19_Case_Death

Death case:
1. convert type of submission_date to date 
2. select date which is in the same range as COVID-19 hospital capacity data-set
3. select useful features
4. add column 'population' by state
5. keep the record with latest date for each state as features 'tot_cases' and 'tot_death' are cumulative features

Dataframes after transformation:
```{r}
case_state <- case_death_selected [case_death_selected $submission_date==end_date_hos,]
summary(case_state )
```

After the selection, there are 52 records with 5 features. Each record represents one state in the United States.

### Merge

In order to see the relationship between hospital capacity and death rate of COVID-19, added new columns which would represents the death rate, total number of confirmed inpatients, hospitalization rate and percentage of total beds in population for each state. 

```{r}
hospital_case<-merge(hospital_state,case_state,by="state")

# remove date column
hospital_case<-hospital_case[,-grep("submission_date",colnames(hospital_case))]

# calculate death rate for each state
hospital_case['death_rate']<-hospital_case$tot_death/hospital_case$tot_cases

# calculate total number of confirmed inpatient for each state
hospital_case['tot_confi_inpatient']<-hospital_case$adu_conf+hospital_case$ped_conf

# calculate the hospitalization rate for each state
hospital_case['inhospi_rate']<-hospital_case$tot_confi_inpatient/hospital_case$tot_cases

# calculate the percentage of number of beds in total population for each state
hospital_case['beds_perc']<-(hospital_case$total_beds+hospital_case$total_icu_beds)/hospital_case$population

summary(hospital_case)
```

So the final data-set after transformation will have 52 observations with 12 features.
















<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

## US vaccination 

### Dataframe after data cleaning
```{r}
head(us_whole_vaccination)
```

The original dataframe contain vaccination information of all states and some organizations of United States. We extracted whole country data and we can see that each row represents a day from 2020-12-20 to now. The information is updated everyday. There are lots of missing values in the table and we are going to use different methods to visualize it.

### Missing value counts by column
```{r,echo=FALSE}
aa <- colSums(is.na(us_whole_vaccination)) %>% sort(decreasing = TRUE)
result <- data.frame(num_missing = aa)
result
```

This is the missing value counts by different features. According to this table, we can see that the feature 
"daily_vaccinations_raw" has the largest number of missing values. And for some features, they have similar number of missing values. Such as "people_fully_vaccinated_per_hundred" and "people_fully_vaccinated", both of them have number of missing values equal to 35, also, according the the features' names, it can be guessed that there is a correlation between this two value, which may be the number of missing value in "people_fully_vaccinated_per_hundred" is dependent on the feature "people_fully_vaccinated", so they always have the missing value at the same time. 


### Missing value counts by row
```{r,echo=FALSE}
values <- rowSums(is.na(us_whole_vaccination))
result <- data.frame(date=as.Date(character()),
                     num_missing=integer())
na_rows_index <- c()
count <- 1
for(i in 1:length(values)){
  if(values[i] != 0){
    result[count,1] <- us_whole_vaccination[i,1]
    result[count,2] <- values[i]
    count <- count + 1
    na_rows_index <- c(na_rows_index,i)
  }
}
result1 <- result
result1$date <- format(result1$date,"%Y-%m")
result1 <- result1 %>% group_by(date) %>% summarise(num_missing = sum(num_missing))
result1
```

Each row in the original dataframe represents a specific day. However, there are too many rows that have missing values. Therefore, we count missing values by month.

December 2020 and January 2021 have the greatest number of missing values. This is reasonable as vaccination just started and it's hard to get all data well-organized.

### Heatmap
```{r,echo=FALSE,fig.height=10,fig.width=10}
df <- us_whole_vaccination[na_rows_index,-2] %>% 
  gather(key,value,-date) %>%
  mutate(missing=ifelse(is.na(value),'yes','no'))

ggplot(df,aes(x=key,y=fct_rev(as.character(date)),fill=missing)) +
  geom_tile(color='white') +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,vjust=0.6)) +
  labs(y='date')
  
```

In order to see the missing patterns we have for this dataframe, we can use heatmap. It can be observed that rows that have missing values all at once by using this heatmap. However, since there are 330 rows in the us_whole_vaccination dataframe, it is hard to summarize the missing value patterns without showing the whole graph. Moreover, as looking through the graph, several rows have the similar missing value patterns. So instead of showing the heatmap for each row, plot just show the different missing value patterns for the dataset is prefered.


```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=10}
source("functions/PSet4b.R")
missing_plot(us_whole_vaccination,FALSE)
```

This missing value plot is a combination of all the information above. It consists of three plots. 

The main plot is the heatmap on the bottom left. It summarized different missing value patterns among all rows. The blue block in each column represents there are missing values in this column. According to the main plot, it shows that there are 7 different patterns for us_whole_vaccination dataset, which are represented by each row of the heatmap. The first row in the heatmap represents the complete cases, which means there is no missing value for rows in this pattern. By looking through this plot, there is a correlation between some features as mentioned in 4.2. For example, for features "people_fully_vaccinated_per_hundred" and "people_fully_vaccinated", in each pattern, their missing value always occurred in the same time. Similar for features "total_distributed, people_vaccinated, people_vaccinated_per_hundred, distributed_per_hundred, and share_doses_used", "daily_vaccinations and daily_vaccinations_per_million", and "total_vaccinations and  total_vaccinations_per_hundred". It indicates that for some of these features, it maybe reasonable to only keep one feature from each group. For example, keep "people_fully_vaccinated" and remove "people_fully_vaccinated_per_hundred".

The barplot on the right is the row count plot. It tells us the number of rows that each pattern can apply. It shows that the complete cases pattern has the most largest number of rows. From this plot, it shows that pattern 2 has the largest number of rows with missing data. By checking with this pattern, it contains 10 features that are missing. A reasonable guess is that the most of data for this pattern comes from December 2020 and January 2021. Crossing checking with the heatmap for each row in 4.4, this guess is confirmed. 

The barplot on the top is a visualization of 4.2. It shows how many rows contain missing values of different features.

As there are not many missing rows, we might remove missing patter 2, 4, 5, and 7 as over 50% of the features are missing. For pattern 3 and 6, we can use mean or median to fill the missing values. 

## COVID-19_Hospital_Capacity

In order to see if there is any state has large number of missing values, missing table was built based on the data set before transformation. Below is the first five rows of the missing-value table. It counts the number of missing value in each feature for each state. According to the table, it indicates that there is no state that has more than half of its records are missing value. Therefore, for total_beds, total_icu_beds, aud_conf and ped_conf, choose the average value of these features in each state to replace the missing value. 

```{r}
#source('data/hospital_preprocessing.R')
```

```{r}
# number of records for each state
state_count<-hospital_capacity_selected %>% group_by(state) %>% count(state)

# build missing table by state
df_na <- data.frame(is.na(hospital_capacity_selected))
df_na$state<-hospital_capacity_selected$state
agg <- aggregate(df_na[-3], df_na['state'],sum)

missing_table<-mutate(agg,state_count)
head(missing_table)


```
## COVID-19_Case_Death

This is the missing value counts by different features for COVID-19_Case_Death data-set. According to this table, we can see that there is no missing value for this data-set.

```{r}
aa <- colSums(is.na(case_death_selected)) %>%  sort(decreasing = TRUE)
result_col <- data.frame(num_missing = aa)
result_col

```





<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

## US time series
```{r,warning=FALSE,message=FALSE}
us_cases <- state_cases %>% group_by(date) %>% summarise(case_grow=sum(case_diff))
us_vacc <- state_vaccination %>% group_by(date) %>% summarise(vacc_grow=sum(vacc_diff))
df <- inner_join(us_cases,us_vacc,by='date')

p1 <- ggplot(df,aes(x=date, y= case_grow)) +
  geom_line(size=1,color='red1',show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Case Growth') +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p2 <- ggplot(df,aes(x=date, y= vacc_grow)) +
  geom_line(size=1,color='springgreen4',show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Vaccinated Growth') +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p1 + p2 + plot_layout(ncol = 1) + plot_annotation(title='US Time Series')
```

This graph shows the time series of confirmed cases growth and people vaccinated growth. It's obvious that when people vaccinated growth number is high, confirmed cases growth speed decreases and when people vaccinated low, confirmed cases increases again. This plot confirms that vaccination helps to prevent covid spread. 

## US cumulative cases
```{r,message=FALSE,warning=FALSE}
us_geo <- tigris::states(class = "sf") 
df <- state_cases[state_cases$date == day,]
shape_obj <- left_join(us_geo, df, by=c('NAME'='state')) %>% 
  filter(REGION != 9) %>% 
  filter(NAME != 'District of Columbia')

us_map <- tm_shape(shape_obj) +
  tm_polygons('cases',id='Name') +
  tm_layout(main.title=paste('Total Cases On ',day),
            main.title.position = "left",
            main.title.size = 1)

dff <- state_cases_raw %>% 
  select(UID:Combined_Key,'12/9/21')

dff_sf <- st_as_sf(dff,coords=c('Long_','Lat'),crs=4326)

tmap_mode('view')

tm_shape(shape_obj) +
  tm_polygons('cases',id='Name') +
  tm_layout(main.title=paste('Total Cases On ',day),
            main.title.position = "left",
            main.title.size = 1,
            legend.outside = TRUE,
            legend.position = c("left", "top")) +  
tm_shape(dff_sf) +
  tm_symbols(col='red',size='12/9/21',scale=1) 
```

This is a map that shows cumulative confirmed cases in United States on 12/9/2021. Red points mean confirmed cases and the size represents exact number. The higher the number, the larger the spot. We can see that large cities are serious spots. This graph gives us a general view of current covid situation.

## Latest cummulative cases and deaths
```{r}
# day <- min(max(state_cases$date),max(state_death$date),max(state_vaccination$date))

ggplot(state_cases[state_cases$date==day,],aes(reorder(state,cases),cases)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Cases',title=paste('Total Cases On ',day)) +
  theme_bw()

ggplot(state_death[state_death$date==day,],aes(reorder(state,death),death)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Death',title=paste('Total Death On ',day)) +
  theme_bw()

# df1 <- state_cases[state_cases$date==day,] %>% mutate(type='cases',value=cases)
# df2 <- state_death[state_death$date==day,] %>% mutate(type='death',value=death)
# df <- rbind(df1,df2)
# 
# df2 <- df %>% filter(type=='cases') %>% arrange(value)
# the_order <- df2$state
# 
# ggplot(df,aes(x = state, y = value, group = type, fill = type)) +
#   geom_bar(stat = "identity", width = 0.75) +
#   coord_flip() +
#   scale_x_discrete(limits = the_order) +
#   scale_y_continuous(breaks = seq(-5000000, 80000, 1000000), 
#                      labels = abs(seq(-5000000, 80000, 1000000))) +
#   labs(x = "State", y = "Count", title = "Back-to-back bar chart") +
#   theme(legend.position = "bottom",
#         legend.title = element_blank(),
#         plot.title = element_text(hjust = 0.5),
#         panel.background = element_rect(fill =  "grey90")) +
#   scale_fill_manual(values=c("red", "blue"),
#                     name="",
#                     breaks=c("Case", "Death"),
#                     labels=c("Case", "Death"))
# 
# ggplot(df, aes(x=state, y= value, fill=type)) + 
#     facet_wrap(~ type, scales = "free_x") + 
#     geom_col() + 
#     coord_flip() +
#     scale_y_continuous(expand = c(0, 0)) +
#     theme(panel.spacing.x = unit(0, "mm"))
```

The graphs above shows current confirmed cases and death of 50 states. We can see that California, Texas, Florida, New York, and Illinois are top 5 states with most confirmed cases. California, Texas, Florida, New York, and Pennsylvania are top 5 states with most deaths. 


## Latest vaccination situation
```{r,fig.width=9,fig.height=9,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases[,-4],by=c('date','state')),state_death[,-4])
df <- df[,c(-1,-2)]
cormat <- round(cor(df),2)

melted_cormat <- melt(cormat)

ggplot(data=melted_cormat,aes(Var1,Var2,fill=value))+
  geom_tile() +
  geom_text(aes(label=value)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = 'Correlation Heat Map of Vaccination Features')
```

As there are 12 features in our vaccination data, we use a correlation heatmap to select the most relevant features to confirmed cases and deaths. From the graph above we can see the top 4 features are total vaccinations, total distributed, people vaccinated, and people fully vaccinated. We will use people vaccinated and people fully vaccinated as two features to represent vaccination.

```{r}
df <- state_vaccination[,c(1,2,5,6)]
p1 <- ggplot(df[df$date == day,],aes(reorder(state, people_vaccinated),people_vaccinated)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Vaccinated') +
  theme_bw()

p2 <- ggplot(df[df$date == day,],aes(reorder(state, people_fully_vaccinated_per_hundred),people_fully_vaccinated_per_hundred)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Fully Vaccinated Per Hundred') +
  theme_bw()

p1 + p2 + plot_annotation(title=paste('Vaccination Situation On ',day))
```

The 5 states with most deaths also have the largest number of people vaccinated. However, none of them are on top of the list of people fully vaccinated.

## Time series of serious states
```{r,warning=FALSE,message=FALSE}
states <- c('California', 'Texas', 'Florida', 'New York', 'Pennsylvania')

df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(state %in% states) 
df <- subset(df,vacc_diff < 500000)
  
p1 <- ggplot(df, aes(x=date, y= case_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme(legend.position = "none") +
  labs(y='Case') +
  theme_minimal() +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p2 <- ggplot(df, aes(x=date, y= death_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme(legend.position = "none") +
  labs(y='Death') +
  theme_minimal() +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p3 <- ggplot(df, aes(x=date, y= vacc_diff)) +
  geom_line(aes(color=state),size=1) +
  labs(y='Vaccinated') +
  theme_minimal() +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p4 <- ggplot(df, aes(x=date, y= fully_vacc_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Fully Vaccinated') +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

p1 + p2 + p3 + p4 + plot_layout(ncol = 1) + plot_annotation(title='Growth Time Series')
```

There is a significant outlier in original people_vaccinated daily growth graph. Because from October 2nd to November 28th, the people_vaccinated information in Pennsylvania is missing. We are doing a filling by pervious data, therefore there exists a huge pike. We have to delete the data of that day.

## Distribution
```{r,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(date == day)

ggplot(df, aes(cases, people_fully_vaccinated, color=state)) +
  geom_point(show.legend = FALSE) +
  geom_smooth(method=lm,se=FALSE) +
  geom_text(aes(label=state),size=2.6,hjust=0,vjust=0) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x='Cases',y='People Fully Vaccinated',title=paste('Vaccination vs. Cases Scatterplot On ',day))
  
```

This scatterplot shows cumulative people fully vaccinated vs. cases in all states. We can easily see the outliers among 50 states. If we see in ratio, there's no significantly high or low states.

## Percentage of number of beds in population VS death rate
```{r}

ggplot(hospital_case, aes(beds_perc, death_rate)) + 
  geom_point(size=0.8,alpha=0.7) +
  geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
  theme_bw() +
  labs(x='percentage of beds in population',y='death rate',title='Scatter Plot for beds VS death rate ')

```

According to the graph, there is no clear trend between the percentage of beds inn population and the death rate. Especially when we ignore the state which has more than 0.035. It’s different from what we initially imagined, when capacity of hospital in a state increase, the mortality rate will drop. It may cause by several reasons. The first reason may be due to inaccurate mortality. Because the symptoms of COVID-19 are similar to those of flu, many people will mistakenly think that they are flu instead of COVID-19, leading to deviations in statistical data. The second possibility is that there are other factors, such as the number of ventilators or whether the isolation measures are effective. These are also factors that can affect mortality. 

## inpatient rate and death rate

```{r}
ggplot(hospital_case, aes(inhospi_rate, death_rate)) + 
  geom_point(size=1) +
  geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
  theme_bw() +
  labs(x='inpatient rate',y='death rate',title='Scatter Plot for inpatient rate VS death rate ')

```

Form the above graph, we can see that death rate increase as the inpatient rate increases. There are two possible explanations. The first one is people would not go to hospital unless they are in very bad situation. Thus as the number of impatient increase, death rate increase. The second explanation is consider about the time range. There is no very useful way to cure COVID-19 patient. Thus as inpatient number increase, death rate increase. 









<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component

## Top 5 series states
<script src="https://d3js.org/d3.v4.js"></script>

<!-- Add 2 buttons -->
<button onclick="update('cases')">Cases</button>
<button onclick="update('death')">Deaths</button>
<button onclick="update('people_fully_vaccinated')">People Fully Vaccinated</button>

<!-- Create a div where the graph will take place -->
<div id="my_dataviz"></div>

<script>

// set the dimensions and margins of the graph
var margin = {top: 30, right: 30, bottom: 70, left: 60},
    width = 460 - margin.left - margin.right,
    height = 400 - margin.top - margin.bottom;

// append the svg object to the body of the page
var svg = d3.select("#my_dataviz")
  .append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)
  .append("g")
    .attr("transform",
          "translate(" + margin.left + "," + margin.top + ")");

// Initialize the X axis
var x = d3.scaleBand()
  .range([ 0, width ])
  .padding(0.2);
var xAxis = svg.append("g")
  .attr("transform", "translate(0," + height + ")")

// Initialize the Y axis
var y = d3.scaleLinear()
  .range([ height, 0]);
var yAxis = svg.append("g")
  .attr("class", "myYaxis")


// A function that create / update the plot for a given variable:
function update(selectedVar) {

  // Parse the Data
  d3.csv('https://raw.githubusercontent.com/yrz1997/STAT5702-Final-Project-Group-7/main/data/d3_data.csv', function(data) {

    // X axis
    x.domain(data.map(function(d) { return d.state; }))
    xAxis.transition().duration(1000).call(d3.axisBottom(x))

    // Add Y axis
    y.domain([0, d3.max(data, function(d) { return +d[selectedVar] }) ]);
    yAxis.transition().duration(1000).call(d3.axisLeft(y));

    // variable u: map data to existing bars
    var u = svg.selectAll("rect")
      .data(data)

    // update bars
    u
      .enter()
      .append("rect")
      .merge(u)
      .transition()
      .duration(1000)
        .attr("x", function(d) { return x(d.state); })
        .attr("y", function(d) { return y(d[selectedVar]); })
        .attr("width", x.bandwidth())
        .attr("height", function(d) { return height - y(d[selectedVar]); })
        .attr("fill", "#69b3a2")
  })

}

// Initialize plot
update('cases')

</script>

This interactive graph shows total cases, deaths, and people fully vaccinated numbers on 2021-12-9. Compared to static graphs, it's easier for us to visualize the difference between these properties.


<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

