--- 
title: "United States Covid Project"
author: "Ruize Yu, Jiazhen Li, Wenxin Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
always_allow_html: yes
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction




<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

In this project, we have three data sources

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Vaccination data
We collected vaccination data from an organization called 'Our World in Data'. They gather vaccination information of United States from official and update the data in a daily basis. The daily cases and death data are gathered from Center for Systems Science and Engineering at John Hopkins University. It's updated every 24 hours as well. 

For the vaccination data, there are 14 features. The names of them are listed below:
```{r}
print(colnames(state_vaccination_raw))
```

The starting date of this dataset is 2021-01-12 and the last day is yesterday.
The first problem we noticed was that in location, there are many places that are not US states. Like "Bureau of Prisons","Dept of Defense"... Therefore, the first step that we did was to filter out these none-states. Then we found a common issue, missing data. We visualized it in missing values chapter and decided to do a backward filling. Because most of these missing values were not continuous, we could simply use the latest values to fill in these missing spots.

For the cases and death data, the raw data put every day as a column therefore we first converted these columns to rows.
```{r}
print(colnames(state_cases_raw)[c(1:15)])
```

The starting date of this dataset is 2020-01-22 and the last day is yesterday.
When we tried to calculate daily growing cases and death from cummulative data, we noticed that there were mistakes when collecting these data because some of the growing numbers were negative. We used the same filling method as in vaccination data.



## COVID-19-Reported-Patient-Impact-and-Hospital-Capa

```{r,message=FALSE,warning=FALSE,echo=FALSE}
source('data/hospital_preprocessing.R')
```

Source: https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u/data

This data-set provides facility-level data for hospital utilization aggregated on a weekly basis (Friday to Thursday) in the time period from 2020-07-31 to 2021-11-26. It is mainly collected from two sources: (1) HHS TeleTracking, and (2) reporting provided directly to HHS Protect by state/territorial health departments on behalf of their healthcare facilities.

There are 106 features with 309719 records in the original data-set, but not all of those features are useful for our analysis, so we only select 8 specific features from 106 original features. These selected features including date, state, capacities of hospitals and number of hospitalized patients who are confirmed with COVID-19. 

We also noticed that there is value -999999 in data-set which is not a reasonable value. Replacing -999999 with 2, since it represents sums and averages less than four, so use average to replace that.

```{r}
#summary of data-set

summary(hospital_capacity_selected)

```

## United-States-COVID-19-Cases-and-Deaths-by-State

Source: https://data.cdc.gov/Case-Surveillance/United-States-COVID-19-Cases-and-Deaths-by-State-o/9mfq-cb36/data

This data-set is collected by CDC, which includes counts of COVID-19 cases and death numbers daily in the time period form 2020-01-22 to 2021-12-07. CDC collects and organizes these data based on these most recent numbers reported by states, territories, and other jurisdictions. There are 41160 records with 15 features in the original data-set, and each record represents a specific date for a specific state. We only keep 3 features which are state, total number of cases and total number of death for our further analysis.

Because we want to merge this data-set with 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa', so it is important to make sure they are in the same time period. Since the time range for 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' is smaller, so the subset of original data-set is generated with the time period from 2020-07-31 to 2021-11-26. The sub data-set has 29040 records with 3 features. Note that for the features 'tot_cases' and 'tot_death', they are cumulative features during the time period.

```{r}
#summary of data-set

summary(case_death_selected)

```















<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Vaccination, case, and death data
The code is in data/data_preprocessing.R

For this part, there are 3 datasets: vaccination, cases, and death. Cases and death are from same source, vaccination is from another source.

Vaccination:
1. Filter out non-state, only keep 50 states. 
2. Fill NA with previous value.
3. Rename 'location' to 'state' for later join with other dataframe. Change 'New York State' to 'New York' to be consistent with other data source.
4. When calculating growth, fill first date with 0 otherwise it would be NA.

Dataframe after transformation:
```{r}
head(state_vaccination)
```

Case and death:
1. Filter out non-state, only keep 50 states.
2. Convert date columns to rows by using dplyr.
3. Change date from character type to date type.
4. Rename 'Province_state' to 'state' for later join with other dataframe.
5. When calculating growth, fill first date with 0 otherwise it would be NA. Change all negative values to NA, then fill with previous value. There shouldn't have negative values in growth because cumulative covid cases and deaths shouldn't decrease as time goes by.

Dataframes after transformation:
```{r}
head(state_cases)

head(state_death)
```

## Relationship between morbidity and mortality and hospital capacity
In order to do see the relationship, two data-set 'COVID-19-Reported-Patient-Impact-and-Hospital-Capa' and 'United-States-COVID-19-Cases-and-Deaths-by-State' are used.

### COVID-19_Hospital_Capacity
To see the difference in hospital capacity and number of inpatient who is confirmed with COVID-19 between state, group all hospital records for each state. 

```{r}
source('data/hospital_preprocessing.R')
```

```{r}
# fill NA by using mean value for each feature in each state
impute_mean <- function(x) replace(x, is.na(x), round(mean(x, na.rm = TRUE)))
hospital_capacity_selected <- ddply(hospital_capacity_selected, ~ state, transform, total_beds = impute_mean(total_beds),
     total_icu_beds = impute_mean(total_icu_beds),adu_conf = impute_mean(adu_conf),ped_conf = impute_mean(ped_conf))

hospital_state <- hospital_capacity_selected %>% group_by(state) %>% summarize_if(is.numeric,sum,na.rm = TRUE)

summary(hospital_state)
```

After the transformation, there are 56 observations with 7 features, each record represents one state in the United States.


### COVID-19_Case_Death
To merge it with the Hospital Capacity data-set, records also needed to be grouped by state. However, as mentioned before, as features 'tot_cases' and 'tot_death', they are cumulative features during the time period, so only keep the record with latest date for each state. 

```{r}
case_state <- case_death_selected [case_death_selected $submission_date==end_date_hos,]
summary(case_state )
```

After the selection, there is 60 records with 4 features.

### Merge
In order to see the relationship between hospital capacity and death rate of COVID-19, added new columns which would represents the death rate, total number of confirmed inpatients, hospitalization rate, percentage of total beds in population and percentage of total icu beds in population for each state. 

```{r}
hospital_case<-merge(hospital_state,case_state,by="state")

# remove date column
hospital_case<-hospital_case[,-grep("submission_date",colnames(hospital_case))]

# calculate death rate for each state
hospital_case['death_rate']<-hospital_case$tot_death/hospital_case$tot_cases

# calculate total number of confirmed inpatient for each state
hospital_case['tot_confi_inpatient']<-hospital_case$adu_conf+hospital_case$ped_conf

# calculate the hospitalization rate for each state
hospital_case['inhospi_rate']<-hospital_case$tot_confi_inpatient/hospital_case$tot_cases

# calculate the percentage of number of beds in total population for each state
hospital_case['beds_perc']<-hospital_case$total_beds/hospital_case$population

# calculate the percentage of number of icu beds in total population for each state
hospital_case['icu_beds_perc']<-hospital_case$total_icu_beds/hospital_case$population

summary(hospital_case)
```



















<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

## US vaccination 

```{r,message=FALSE,warning=FALSE,echo=FALSE}
source('data/data_preprocessing.R')
```

### Dataframe after data cleaning
```{r}
head(us_whole_vaccination)
```

The original dataframe contain vaccination information of all states and some organizations of United States. We extracted whole country data and we can see that each row represents a day from 2020-12-20 to now. The information is updated everyday. There are lots of missing values in the table and we are going to use different methods to visualize it.

### Missing value counts by column
```{r,echo=FALSE}
aa <- colSums(is.na(us_whole_vaccination)) %>% sort(decreasing = TRUE)
result <- data.frame(num_missing = aa)
result
```

This is the missing value counts by different features. According to this table, we can see that the feature 
"daily_vaccinations_raw" has the largest number of missing values. And for some features, they have similar number of missing values. Such as "people_fully_vaccinated_per_hundred" and "people_fully_vaccinated", both of them have number of missing values equal to 35, also, according the the features' names, it can be guessed that there is a correlation between this two value, which may be the number of missing value in "people_fully_vaccinated_per_hundred" is dependent on the feature "people_fully_vaccinated", so they always have the missing value at the same time. 


### Missing value counts by row
```{r,echo=FALSE}
values <- rowSums(is.na(us_whole_vaccination))
result <- data.frame(date=as.Date(character()),
                     num_missing=integer())
na_rows_index <- c()
count <- 1
for(i in 1:length(values)){
  if(values[i] != 0){
    result[count,1] <- us_whole_vaccination[i,1]
    result[count,2] <- values[i]
    count <- count + 1
    na_rows_index <- c(na_rows_index,i)
  }
}
result1 <- result
result1$date <- format(result1$date,"%Y-%m")
result1 <- result1 %>% group_by(date) %>% summarise(num_missing = sum(num_missing))
result1
```

Each row in the original dataframe represents a specific day. However, there are too many rows that have missing values. Therefore, we count missing values by month.

December 2020 and January 2021 have the greatest number of missing values. This is reasonable as vaccination just started and it's hard to get all data well-organized.

### Heatmap
```{r,echo=FALSE,fig.height=10,fig.width=10}
df <- us_whole_vaccination[na_rows_index,-2] %>% 
  gather(key,value,-date) %>%
  mutate(missing=ifelse(is.na(value),'yes','no'))

ggplot(df,aes(x=key,y=fct_rev(as.character(date)),fill=missing)) +
  geom_tile(color='white') +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45,vjust=0.6)) +
  labs(y='date')
  
```

In order to see the missing patterns we have for this dataframe, we can use heatmap. It can be observed that rows that have missing values all at once by using this heatmap. However, since there are 330 rows in the us_whole_vaccination dataframe, it is hard to summarize the missing value patterns without showing the whole graph. Moreover, as looking through the graph, several rows have the similar missing value patterns. So instead of showing the heatmap for each row, plot just show the different missing value patterns for the dataset is prefered.


```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=10}
source("functions/PSet4b.R")
missing_plot(us_whole_vaccination,FALSE)
```

This missing value plot is a combination of all the information above. It consists of three plots. 

The main plot is the heatmap on the bottom left. It summarized different missing value patterns among all rows. The blue block in each column represents there are missing values in this column. According to the main plot, it shows that there are 7 different patterns for us_whole_vaccination dataset, which are represented by each row of the heatmap. The first row in the heatmap represents the complete cases, which means there is no missing value for rows in this pattern. By looking through this plot, there is a correlation between some features as mentioned in 4.2. For example, for features "people_fully_vaccinated_per_hundred" and "people_fully_vaccinated", in each pattern, their missing value always occurred in the same time. Similar for features "total_distributed, people_vaccinated, people_vaccinated_per_hundred, distributed_per_hundred, and share_doses_used", "daily_vaccinations and daily_vaccinations_per_million", and "total_vaccinations and  total_vaccinations_per_hundred". It indicates that for some of these features, it maybe reasonable to only keep one feature from each group. For example, keep "people_fully_vaccinated" and remove "people_fully_vaccinated_per_hundred".

The barplot on the right is the row count plot. It tells us the number of rows that each pattern can apply. It shows that the complete cases pattern has the most largest number of rows. From this plot, it shows that pattern 2 has the largest number of rows with missing data. By checking with this pattern, it contains 10 features that are missing. A reasonable guess is that the most of data for this pattern comes from December 2020 and January 2021. Crossing checking with the heatmap for each row in 4.4, this guess is confirmed. 

The barplot on the top is a visualization of 4.2. It shows how many rows contain missing values of different features.

As there are not many missing rows, we might remove missing patter 2, 4, 5, and 7 as over 50% of the features are missing. For pattern 3 and 6, we can use mean or median to fill the missing values. 

## COVID-19_Hospital_Capacity

In order to see if there is any state has large number of missing values, missing table was built based on the data set before transformation. Below is the first five rows of the missing-value table. It counts the number of missing value in each feature for each state. According to the table, it indicates that there is no state that has more than half of its records are missing value. Therefore, for total_beds, total_icu_beds, aud_conf and ped_conf, choose the average value of these features in each state to replace the missing value. 

```{r}
source('data/hospital_preprocessing.R')

# number of records for each state
state_count<-hospital_capacity_selected %>% group_by(state) %>% count(state)

# build missing table by state
df_na <- data.frame(is.na(hospital_capacity_selected))
df_na$state<-hospital_capacity_selected$state
agg <- aggregate(df_na[-2], df_na['state'],sum)

missing_table<-mutate(agg,state_count)
head(missing_table)


```
## COVID-19_Case_Death

This is the missing value counts by different features for COVID-19_Case_Death data-set. According to this table, we can see that there is no missing value for this data-set.

```{r}
aa <- colSums(is.na(case_death_selected)) %>%  sort(decreasing = TRUE)
result_col <- data.frame(num_missing = aa)
result_col

```







<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

```{r,message=FALSE,warning=FALSE}
source('data/data_preprocessing.R')
```

## Latest cummulative cases and deaths
```{r}
day <- min(max(state_cases$date),max(state_death$date),max(state_vaccination$date))

ggplot(state_cases[state_cases$date==day,],aes(reorder(state,cases),cases)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Cases',title=paste('Total Cases On ',day)) +
  theme_bw()

ggplot(state_death[state_death$date==day,],aes(reorder(state,death),death)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='Total Death',title=paste('Total Death On ',day)) +
  theme_bw()
```

The graphs above shows current confirmed cases and death of 50 states. We can see that California, Texas, Florida, New York, and Illinois are top 5 states with most confirmed cases. California, Texas, Florida, New York, and Pennsylvania are top 5 states with most deaths. 


## Latest vaccination situation
```{r,fig.width=9,fig.height=9,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases[,-4],by=c('date','state')),state_death[,-4])
df <- df[,c(-1,-2)]
cormat <- round(cor(df),2)

melted_cormat <- melt(cormat)

ggplot(data=melted_cormat,aes(Var1,Var2,fill=value))+
  geom_tile() +
  geom_text(aes(label=value)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = 'Correlation Heat Map of Vaccination Features')
```

As there are 12 features in our vaccination data, we use a correlation heatmap to select the most relevant features to confirmed cases and deaths. From the graph above we can see the top 4 features are total vaccinations, total distributed, people vaccinated, and people fully vaccinated. We will use people vaccinated and people fully vaccinated as two features to represent vaccination.

```{r}
df <- state_vaccination[,c(1,2,5,6)]
p1 <- ggplot(df[df$date == day,],aes(reorder(state, people_vaccinated),people_vaccinated)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Vaccinated') +
  theme_bw()

p2 <- ggplot(df[df$date == day,],aes(reorder(state, people_fully_vaccinated_per_hundred),people_fully_vaccinated_per_hundred)) +
  geom_bar(stat='identity',color='black',fill='skyblue') +
  coord_flip() +
  labs(x='State',y='People Fully Vaccinated Per Hundred') +
  theme_bw()

p1 + p2 + plot_annotation(title=paste('Vaccination Situation On ',day))
```

The 5 states with most deaths also have the largest number of people vaccinated. However, none of them are on top of the list of people fully vaccinated.

## Time series of serious states
```{r,warning=FALSE,message=FALSE}
states <- c('California', 'Texas', 'Florida', 'New York', 'Pennsylvania')

df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(state %in% states) 
  
p1 <- ggplot(df, aes(x=date, y= case_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Case Growth')

p2 <- ggplot(df, aes(x=date, y= death_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Death Growth')

p3 <- ggplot(df, aes(x=date, y= vacc_diff)) +
  geom_line(aes(color=state),size=1) +
  theme_minimal() +
  labs(y='Vacc Growth')

p4 <- ggplot(df, aes(x=date, y= fully_vacc_diff)) +
  geom_line(aes(color=state),size=1,show.legend = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y='Fully Vacc Growth')

p1 + p2 + p3 + p4 + plot_layout(ncol = 1) + plot_annotation(title='Time Series')
```

There is a significant outlier in people_vaccinated daily growth graph. Because from October 2nd to November 28th, the people_vaccinated information in Pennsylvania is missing. We are doing a filling by pervious data, therefore the huge spike is false.

## Distribution
```{r,warning=FALSE,message=FALSE}
df <- inner_join(inner_join(state_vaccination,state_cases,by=c('date','state')),state_death) %>%
  select(date,state,people_vaccinated,vacc_diff,people_fully_vaccinated,fully_vacc_diff,cases,case_diff,death,death_diff) %>%
  filter(date == day)

ggplot(df, aes(cases, people_fully_vaccinated, color=state)) +
  geom_point(show.legend = FALSE) +
  geom_text(aes(label=state),size=2.6,hjust=0,vjust=0) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x='Cases',y='People Fully Vaccinated',title=paste('Vaccination vs. Cases Scatterplot On ',day))
  
```



## percentage of number of beds in population and death rate
```{r}
ggplot(hospital_case, aes(beds_perc, death_rate)) + geom_point(size=1) +geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
ggtitle("Scatter Plot for beds VS death rate")

```
No clear trend occurs. It may indicates that there is no strong relationship between percentage of number of beds in population and death rate.


## percentage of number of icu beds in population and death rate

```{r}
ggplot(hospital_case, aes(icu_beds_perc, death_rate)) + geom_point(size=1) +geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
ggtitle("Scatter Plot for icu beds VS death rate")

```
There is a more clear trend between total number of icu hospital beds and death rate. As number of beds increase, there is an increase trend in the value of death rate.

## inpatient rate and death rate

```{r}
ggplot(hospital_case, aes(hospi_rate, death_rate)) + geom_point(size=1) +geom_smooth(method = "lm",formula= y~x, se = FALSE,alpha = .05,size=0.4)+
ggtitle("Scatter Plot for tinpatient rate VS death rate")

```

Death rate increase as the inpatient rate. Two possible explanations. The first one is people would not go to hospital unless they are in very bad situation. Thus as the number of impatient increase, death rate increase. The second explanation is consider about the time range. There is no very useful way to cure COVID-19 patient. Thus as inpatient number increase, death rate increase. 









<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component



<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

